---
title: "fongping hydro main"
format: pdf
editor: visual
---

# SETUP

## PACKAGE

```{r}
packages <- c(
  "tidyverse",
  "here",
  "ggplot2",
  "httr2",
  "jsonlite"
)

installed_packages <- rownames(installed.packages())

for (pkg in packages) {
  if (!pkg %in% installed_packages) {
    install.packages(pkg)
  }
}

invisible(lapply(packages, library, character.only = TRUE))
```

## MODULES

```{r}
# source modules
source(here("module", "hydro_reservoir.R"))
source(here("module", "hydro_sediment.R"))
source(here("module", "hydro_finance.R"))
source(here("module", "hydro_climate.R"))
source(here("module", "hydro_test.R"))

## Run tests first — stop if anything fails
run_all_tests()
```

## Load data

```{r}
## Load clean data
## data is prehandled by scripts in data_handle folder

daily_flow    <- read_csv(here("data", "lishan_daily_clean.csv"))
decadal_flow  <- read_csv(here("data", "lishan_decadal_long.csv"))
sed_clean     <- read_csv(here("data", "sediment_clean.csv"))
weir_design   <- read_csv(here("data", "weir_design_flow.csv"))
```





## Analysis starts here...```

## READ DATA

```{r}
# read here("data/ # 10dayrunning"),

ls()
```

```{r}
source(here("module", "hydro_reservoir.R"))

daily_flow <- read_csv(here("data", "lishan_daily.csv"),
                       show_col_types = FALSE)

# Quick test: one year, recommended e-flow
sim_test <- run_reservoir_simulation(
  daily_flow  = daily_flow |> filter(lubridate::year(date) == 2023),
  e_flow_mode = "recommended"
)

# Check outputs
glimpse(sim_test$combined)
cat("Annual energy W1 (GWh):", sum(sim_test$W1$energy_kwh) / 1e6, "\n")
cat("Annual energy W2 (GWh):", sum(sim_test$W2$energy_kwh) / 1e6, "\n")
cat("E-flow satisfied W1 (%):",
    round(mean(sim_test$W1$eflow_satisfied) * 100, 1), "%\n")
```

```{r}
library(tidyverse)
library(readxl)
library(here)
library(lubridate)

# ── 1. 立山站日流量 ──────────────────────────────────────────
daily_flow <- read_csv(here("data", "lishan_daily.csv"),
                       show_col_types = FALSE)

cat("=== lishan_daily.csv ===\n")
cat("Rows          :", nrow(daily_flow), "\n")
cat("Date range    :", as.character(range(daily_flow$date)), "\n")
cat("NA count      :", sum(is.na(daily_flow$Q_cms)), "\n")
cat("NA pct        :", round(mean(is.na(daily_flow$Q_cms))*100,1), "%\n")
cat("Max Q (cms)   :", max(daily_flow$Q_cms, na.rm=TRUE), "\n\n")

# ── 2. 旬流量（原始作業資料）────────────────────────────────
decadal_flow <- read_csv(here("data", "lishan_flow.csv"),
                         show_col_types = FALSE)

cat("=== lishan_flow.csv ===\n")
cat("Rows (tendays):", nrow(decadal_flow), "\n")
cat("Year range    :", names(decadal_flow)[2], "to",
    names(decadal_flow)[ncol(decadal_flow)], "\n\n")

# ── 3. 土砂資料 ──────────────────────────────────────────────
sed_raw <- read_excel(here("ref", "豐坪溪河川輸砂量統計.xlsx"))

cat("=== 豐坪溪河川輸砂量統計.xlsx ===\n")
cat("Rows          :", nrow(sed_raw), "\n")
cat("Columns       :", names(sed_raw), "\n")
cat("Year range    :", min(sed_raw$year), "to", max(sed_raw$year), "\n")
cat("Q range (cms) :", min(sed_raw$Discharge_CMS), "to",
    max(sed_raw$Discharge_CMS), "\n\n")

# ── 4. 確認 ref 資料夾還有什麼 ──────────────────────────────
cat("=== ref/ folder ===\n")
fs::dir_ls(here("ref")) |> basename() |> print()
```




```{r}
# ── 檢查土砂資料結構 ──────────────────────────────────────────

# 看前20行
sed_raw |> head(20) |> print()

# 看有幾個 sheet
excel_sheets(here("ref", "豐坪溪河川輸砂量統計.xlsx"))

# 看奇怪欄位裡有什麼
cat("\n=== 工作方法欄位的值 ===\n")
sed_raw |> count(`工作方法`) |> print()

cat("\n=== ...9 欄位樣本 ===\n")
sed_raw |> select(`...9`, `...10`) |> 
  filter(!is.na(`...9`)) |> head(10) |> print()

# 看 Q > 500 的紀錄（可能是資料錯誤或真實洪水）
cat("\n=== Q > 500 cms 的紀錄 ===\n")
s
```

```{r}
# 讀工作表3（乾淨資料）
sed_sheet3 <- read_excel(
  here("ref", "豐坪溪河川輸砂量統計.xlsx"),
  sheet = "工作表3"
)

cat("=== 工作表3 ===\n")
cat("Rows    :", nrow(sed_sheet3), "\n")
cat("Columns :", names(sed_sheet3), "\n\n")

glimpse(sed_sheet3)

cat("\n=== 前10行 ===\n")
print(head(sed_sheet3, 10))

cat("\n=== Q 分布 ===\n")
cat("Min Q   :", min(sed_sheet3$Discharge_CMS, na.rm = TRUE), "\n")
cat("Max Q   :", max(sed_sheet3$Discharge_CMS, na.rm = TRUE), "\n")
cat("Q > 500 :", sum(sed_sheet3$Discharge_CMS > 500, na.rm = TRUE), "rows\n")
cat("ppm = 0 :", sum(sed_sheet3$Sed_Con_ppm == 0, na.rm = TRUE), "rows\n")
cat("NA rows :", sum(is.na(sed_sheet3$Discharge_CMS)), "\n")

cat("\n=== 年份分布 ===\n")
sed_sheet3 |> count(year) |> print(n = Inf)
```


```{r}
library(tidyverse)
library(readxl)
library(here)
library(lubridate)
library(zoo)

# ── 讀入已有的乾淨日流量 ─────────────────────────────────────
daily_clean <- read_csv(here("data", "lishan_daily.csv"),
                        show_col_types = FALSE)

# ── BLOCK 3: 重建 sed_fixed ──────────────────────────────────
sed_raw <- read_excel(
  here("ref", "豐坪溪河川輸砂量統計.xlsx"),
  sheet = "工作表3"
)

sed_flagged <- sed_raw |>
  mutate(
    date_full = as.Date(
      sprintf("%04d-%02d-%02d",
              as.integer(year),
              as.integer(month),
              as.integer(date))
    ),
    SSL_expected = Discharge_CMS * Sed_Con_ppm * 0.0864,
    flag = case_when(
      Discharge_CMS > 500 & month %in% c(1,2,3,4,5,11,12) ~ "Q_typo",
      Discharge_CMS > 800  ~ "Q_extreme_check",
      Sed_Con_ppm == 0     ~ "below_detection",
      Sed_Con_ppm > 0 &
        abs(Sus_Load_MTDay - SSL_expected) /
        (SSL_expected + 0.001) > 0.15 ~ "ssl_inconsistent",
      TRUE ~ "ok"
    )
  )

sed_fixed <- sed_flagged |>
  mutate(
    Discharge_CMS = if_else(flag == "Q_typo",
                             Discharge_CMS / 100,
                             Discharge_CMS),
    flag = if_else(flag == "Q_typo", "typo_corrected", flag)
  )

cat("sed_fixed built:", nrow(sed_fixed), "rows\n")
cat("Flag distribution:\n")
sed_fixed |> count(flag) |> print()

# ── Cross-validation ─────────────────────────────────────────
sed_with_date <- sed_fixed |>
  mutate(
    date_full = as.Date(
      sprintf("%04d-%02d-%02d",
              as.integer(year),
              as.integer(month),
              as.integer(date))
    )
  )

cross_check <- sed_with_date |>
  filter(flag %in% c("ok", "typo_corrected", "ssl_inconsistent")) |>
  left_join(
    daily_clean |> select(date, Q_daily = Q_cms, Q_raw = Q_cms_raw),
    by = c("date_full" = "date")
  ) |>
  mutate(
    Q_diff_pct = if_else(
      !is.na(Q_daily) & Q_daily > 0,
      (Discharge_CMS - Q_daily) / Q_daily * 100,
      NA_real_
    ),
    matched = !is.na(Q_raw)   # Q_raw is NA only if date not in daily record
  )

cat("\n=== Cross-validation summary ===\n")
cat("Sediment obs used             :", nrow(cross_check), "\n")
cat("Matched to daily flow record  :", sum(cross_check$matched), "\n")
cat("No match in daily record      :", sum(!cross_check$matched), "\n\n")

cat("=== Q difference (sediment vs daily) ===\n")
cross_check |>
  filter(matched) |>
  summarise(
    n             = n(),
    mean_diff_pct = round(mean(Q_diff_pct, na.rm = TRUE), 1),
    median_diff_pct = round(median(Q_diff_pct, na.rm = TRUE), 1),
    sd_diff_pct   = round(sd(Q_diff_pct, na.rm = TRUE), 1),
    within_10pct  = round(mean(abs(Q_diff_pct) < 10,  na.rm = TRUE) * 100, 1),
    within_20pct  = round(mean(abs(Q_diff_pct) < 20,  na.rm = TRUE) * 100, 1)
  ) |>
  print()

cat("\n=== Largest discrepancies (|diff| > 50%) ===\n")
cross_check |>
  filter(matched, abs(Q_diff_pct) > 50) |>
  select(date_full, Q_sed = Discharge_CMS,
         Q_daily, Q_diff_pct, flag) |>
  arrange(desc(abs(Q_diff_pct))) |>
  print(n = 20)

cat("\n=== Unmatched years (no daily record) ===\n")
cross_check |>
  filter(!matched) |>
  count(year) |>
  print(n = Inf)
```



# baseline

## reservoir module

### run reservoir for a year (time step)

### sediment ?

## MONEY MODULE

### daily energy generation \* 365

### yearly profit

### 100 NPV

# CC scenario (deviation of the inflow increase / maybe percipitation data)

## lishan station data

### CC saga \<-- manipulation of inflow data
