---
title: "Fengping River Hydropower — Main Analysis"
format: pdf
editor: visual
---

# OVERVIEW

This program evaluates the financial feasibility of the proposed Fengping River hydropower project under historic flows and two climate change scenarios (SSP2-4.5 and SSP5-8.5).

**Program contract**

-   **Inputs:** daily streamflow record (1958–2025), sediment rating-curve data, weir design specifications, climate projection parameters, finance parameters (FIT, CAPEX, LTV, loan rate, equity rate, project lifetime).
-   **Outputs:** annual energy generation (GWh), annual revenue (NTD), 20-year NPV, and an NPV feasibility frontier across the loan-rate × LTV parameter grid, reported for each scenario (baseline, SSP2-4.5, SSP5-8.5).
-   **Parameters:** e-flow mode, sediment toggle, number of bootstrap simulations, projection year, finance parameter ranges.

# MODULES

| Module | Purpose | Reuse |
|------------------------|---------------------------|---------------------|
| `hydro_reservoir.R` | Reservoir simulation: flow to energy | Used 3 times (baseline, SSP2-4.5, SSP5-8.5) |
| `hydro_sediment.R` | Sediment load to trap efficiency to storage loss | 1 time |
| `hydro_climate.R` | Bootstrap climate scenarios from historic flow | Used 2 times (SSP2-4.5, SSP5-8.5) |
| `hydro_finance.R` | Energy to revenue to NPV to feasibility frontier | Used 3 times (one per scenario) |
| `hydro_plot.R` | Standardized plots for each output | many |
| `hydro_test.R` | Unit tests for all modules | 1 time at setup |

# SETUP

## PACKAGES

```{r}
#| message: false
#| warning: false

packages <- c(
  "tidyverse", "here", "ggplot2",
  "patchwork", "scales", "lubridate"
)
installed_packages <- rownames(installed.packages())
for (pkg in packages) {
  if (!pkg %in% installed_packages) install.packages(pkg)
}
invisible(lapply(packages, library, character.only = TRUE))
```

## Loading Modules

```{r}
#| message: false

source(here("module", "hydro_reservoir.R"))
source(here("module", "hydro_sediment.R"))
source(here("module", "hydro_finance.R"))
source(here("module", "hydro_climate.R"))
source(here("module", "hydro_plot.R"))
source(here("module", "hydro_test.R"))

# Run all tests first — stops execution if anything fails
run_all_tests()
```

## Loading Data

```{r}
#| message: false

daily_flow   <- read_csv(here("data", "lishan_daily_clean.csv"),
                         show_col_types = FALSE)
decadal_flow <- read_csv(here("data", "lishan_decadal_long.csv"),
                         show_col_types = FALSE)
sed_clean    <- read_csv(here("data", "sediment_clean.csv"),
                         show_col_types = FALSE)
weir_design  <- read_csv(here("data", "weir_design_flow.csv"),
                         show_col_types = FALSE)
rc_params    <- read_csv(here("data", "rating_curve_params.csv"),
                         show_col_types = FALSE)
```

## Financial Parameters - adjust to explore different scenarios

```{r}
# Feed-In Tariff (NTD / kWh)
# Source: 經濟部能源署 114年度再生能源電能躉購費率 (> 20,000 kW class)
FIT_NTD_PER_KWH <- 2.8599
# Total CAPEX (NTD)
# Source: Yongwei investor conference 2025 (up from NTD 6.4B due to inflation)
CAPEX_NTD <- 9.7e9
# Loan-to-value ratio (fraction)
# Typical range for Taiwan utility-scale renewables: 0.70 – 0.80
LTV <- 0.75
# Annual loan interest rate (fraction)
# Current Taiwan syndicated loan range: 0.01 – 0.05
R_LOAN <- 0.03
# Equity discount rate (fraction)
# Required return for private utility investor
R_EQUITY <- 0.08
# Project lifetime (years)
# Source: Shihfeng Power Co. EIA documents
T_YEARS <- 35L

# Sensitivity grid ranges
# Adjust these to zoom in / out on the feasibility frontier
R_LOAN_RANGE <- seq(0.01, 0.05, by = 0.005)   # 1% to 5%
LTV_RANGE    <- seq(0.50, 0.90, by = 0.05)    # 50% to 90%

cat("=== Finance parameters ===\n")
cat(sprintf("FIT rate     : NTD %.4f / kWh\n", FIT_NTD_PER_KWH))
cat(sprintf("CAPEX        : NTD %.1f billion\n", CAPEX_NTD / 1e9))
cat(sprintf("LTV          : %.0f%%\n", LTV * 100))
cat(sprintf("Loan rate    : %.1f%%\n", R_LOAN * 100))
cat(sprintf("Equity rate  : %.1f%%\n", R_EQUITY * 100))
cat(sprintf("Project life : %d years\n", T_YEARS))
```

# Baseline (historic) scenarios

## Reservoir simulation

```{r}
# Run reservoir simulation — baseline, recommended e-flow
sim_baseline <- run_reservoir_simulation(
  daily_flow  = daily_flow,
  e_flow_mode = "recommended"
)
cat("=== Baseline simulation (1958–2025, recommended e-flow) ===\n")
cat("Annual energy W1 (GWh)  :",
    round(sum(sim_baseline$W1$energy_kwh) / 1e6, 1), "\n")
cat("Annual energy W2 (GWh)  :",
    round(sum(sim_baseline$W2$energy_kwh) / 1e6, 1), "\n")
cat("E-flow satisfied W1 (%) :",
    round(mean(sim_baseline$W1$eflow_satisfied) * 100, 1), "%\n")
cat("E-flow satisfied W2 (%) :",
    round(mean(sim_baseline$W2$eflow_satisfied) * 100, 1), "%\n")
```

## Reservoir Operation Plots

```{r}
# Default: first 90 days
plot_reservoir_operation(
  sim_weir    = sim_baseline$W1,
  weir_label  = "W1",
  e_flow_mode = "recommended"
)
# Typhoon season: July–September 2023
plot_reservoir_operation(
  sim_weir    = sim_baseline$W1,
  date_start  = "2023-07-01",
  date_end    = "2023-09-30",
  weir_label  = "W1",
  e_flow_mode = "recommended"
)
# Dry season: January–March 2023
plot_reservoir_operation(
  sim_weir    = sim_baseline$W1,
  date_start  = "2023-01-01",
  date_end    = "2023-03-31",
  weir_label  = "W1",
  e_flow_mode = "recommended"
)
```

## Sediment

```{r}
sed_out <- build_sediment_outputs(
  sed_csv_path = here("data", "sediment_clean.csv"),
  daily_flow   = daily_flow,
  plot_fit     = FALSE
)

cat("\n=== Sediment summary ===\n")
cat("Annual SSL (MT/year):", round(sed_out$annual_ssl_mt), "\n")
cat("\nTrap efficiency W1 (year 1, 10, 35):\n")
sed_out$trap_W1 |>
  filter(year %in% c(1, 10, 35)) |>
  select(year, S_remaining_m3, TE) |>
  print()
```

## Energy Generation Calendar

```{r}
# 單一年份
plot_energy_calendar_heatmap(sim_baseline$W1, year_filter = 2010)
# 某段區間
plot_energy_calendar_heatmap(sim_baseline$W1, year_filter = 2010:2020)
# 全部年份
plot_energy_calendar_heatmap(sim_baseline$W1)
```

## Finance — baseline without sediment

```{r}
total_gwh_baseline <- (sum(sim_baseline$W1$energy_kwh) +
                       sum(sim_baseline$W2$energy_kwh)) / 1e6

fin_no_sed <- build_finance_outputs(
  annual_gwh     = total_gwh_baseline,
  scenario_label = "Baseline (no sediment)"
)
```

## Finance — baseline with sediment adjustment

```{r}
# Sediment reduces storage → reduces power output over time
# Use mean TE over project life to adjust annual GWh
mean_te_W1 <- mean(sed_out$trap_W1$TE)
mean_te_W2 <- mean(sed_out$trap_W2$TE)

# Approximate: energy loss proportional to storage loss
# (simplified — full dynamic coupling in research version)
storage_loss_factor <- 1 - (mean_te_W1 * 0.6 + mean_te_W2 * 0.4)
gwh_with_sed <- total_gwh_baseline * storage_loss_factor

fin_with_sed <- build_finance_outputs(
  annual_gwh     = gwh_with_sed,
  scenario_label = "Baseline (with sediment)"
)

cat("=== Finance comparison ===\n")
cat(sprintf("Without sediment: NTD %.2fB/yr | NPV NTD %.2fB\n",
            fin_no_sed$annual_revenue_ntd / 1e9,
            fin_no_sed$npv_ntd / 1e9))
cat(sprintf("With sediment   : NTD %.2fB/yr | NPV NTD %.2fB\n",
            fin_with_sed$annual_revenue_ntd / 1e9,
            fin_with_sed$npv_ntd / 1e9))
```

## Baseline Feasibility Frontier

```{r}
#| fig-width: 9
#| fig-height: 5

# Sensitivity for baseline scenario
sens_no_sed   <- fin_no_sed$sensitivity
sens_with_sed <- fin_with_sed$sensitivity

sens_combined <- bind_rows(sens_no_sed, sens_with_sed)
user_point <- data.frame(
  scenario   = c("Baseline (no sediment)", "Baseline (with sediment)"),
  r_loan_pct = R_LOAN * 100,
  ltv_pct    = LTV * 100
)

plot_npv_feasibility_frontier(sens_combined) +
  geom_point(
    data   = user_point,
    aes(x = r_loan_pct, y = ltv_pct),
    shape  = 23,
    size   = 4,
    fill   = "white",
    colour = "black",
    stroke = 1.2
  ) +
  labs(caption = paste0(
    "Diamond = current parameter setting",
    " (LTV = ", LTV * 100, "%, rate = ", R_LOAN * 100, "%)\n",
    "CAPEX = NTD ", CAPEX_NTD / 1e9, "B; ",
    "FIT = NTD ", FIT_NTD_PER_KWH, "/kWh; ",
    "Equity rate = ", R_EQUITY * 100, "%"
  ))
```

# CLIMATE CHANGE SCENARIOS

## Generate scenarios

```{r}
#| message: false
scenario_names   <- names(climate_out)
scenario_finance <- vector("list", length(scenario_names))

for (i in seq_along(scenario_names)) {

  ssp_name <- scenario_names[[i]]
  scen     <- climate_out[[ssp_name]]

  # P50 mean flow → rough annual GWh
  p50_gwh <- scen$summary$percentiles |>
    filter(variable == "mean_Q") |>
    pull(P50) * 365 * 86400 / 1e9

  fin <- build_finance_outputs(
    annual_gwh     = p50_gwh,
    scenario_label = ssp_name
  )

  scenario_finance[[i]] <- data.frame(
    scenario   = ssp_name,
    gwh_p50    = p50_gwh,
    npv_b_ntd  = fin$npv_ntd  / 1e9,
    irr_pct    = fin$irr_pct,
    payback_yr = fin$payback_yr
  )
}

scenario_finance <- bind_rows(scenario_finance)
print(scenario_finance)
```

## NPV across scenarios (loop over SSPs)

This is the program's main `for` loop: for each climate scenario, run finance on the P50 mean-flow projection and collect outputs.

```{r}
# Run finance for each scenario P50
scenario_finance <- purrr::imap_dfr(climate_out, function(scen, ssp_name) {

  p50_gwh <- scen$summary$percentiles |>
    filter(variable == "mean_Q") |>
    pull(P50) * 365 * 86400 / 1e9  # rough GWh from mean Q

  fin <- build_finance_outputs(
    annual_gwh     = p50_gwh,
    scenario_label = ssp_name
  )

  data.frame(
    scenario   = ssp_name,
    gwh_p50    = p50_gwh,
    npv_b_ntd  = fin$npv_ntd / 1e9,
    irr_pct    = fin$irr_pct,
    payback_yr = fin$payback_yr
  )
})

print(scenario_finance)
```

## Feasibility Frontier - Climate Scenarios

```{r}
#| fig-width: 9
#| fig-height: 5

sens_no_sed_clim <- run_sensitivity(
  annual_revenue_ntd = fin_no_sed$annual_revenue_ntd,
  r_loan_seq         = R_LOAN_RANGE,
  ltv_seq            = LTV_RANGE,
  scenario_label     = "Without sediment"
)

sens_with_sed_clim <- run_sensitivity(
  annual_revenue_ntd = fin_with_sed$annual_revenue_ntd,
  r_loan_seq         = R_LOAN_RANGE,
  ltv_seq            = LTV_RANGE,
  scenario_label     = "With sediment"
)

sens_combined_clim <- bind_rows(sens_no_sed_clim, sens_with_sed_clim)

user_point_clim <- data.frame(
  scenario_label = c("Without sediment", "With sediment"),
  r_loan_pct     = R_LOAN * 100,
  ltv_pct        = LTV * 100
)

plot_npv_feasibility_frontier(sens_combined_clim) +
  geom_point(
    data        = user_point_clim,
    aes(x = r_loan_pct, y = ltv_pct),
    shape       = 23,
    size        = 4,
    fill        = "white",
    colour      = "black",
    stroke      = 1.2,
    inherit.aes = FALSE
  ) +
  labs(caption = paste0(
    "Diamond = current parameter setting",
    " (LTV = ", LTV * 100, "%, rate = ", R_LOAN * 100, "%)\n",
    "CAPEX = NTD ", CAPEX_NTD / 1e9, "B; ",
    "FIT = NTD ", FIT_NTD_PER_KWH, "/kWh; ",
    "Equity rate = ", R_EQUITY * 100, "%"
  ))
```

# OPERATIONS COMPARISON

## Forecast-informed simulation (24-hour lookahead)

```{r}
sim_forward24 <- run_reservoir_simulation_forward24(
  daily_flow  = daily_flow,
  e_flow_mode = "recommended",
  scale_w2    = TRUE
)

annual_forward24 <- summarise_annual_generation(sim_forward24)

result <- annual_forward24 |>
  summarise(
    mean_energy_gwh    = mean(energy_gwh_total, na.rm = TRUE),
    p10_energy_gwh     = quantile(energy_gwh_total, 0.10, na.rm = TRUE),
    p90_energy_gwh     = quantile(energy_gwh_total, 0.90, na.rm = TRUE),
    mean_revenue_b_ntd = mean(revenue_ntd, na.rm = TRUE) / 1e9,
    mean_spill_m3      = mean(spill_m3_W1 + spill_m3_W2, na.rm = TRUE)
  )
print(result)
```

### Finance — baseline without sediment

```{r}
# Finance — without sediment
fin_no_sed <- build_finance_outputs(
  annual_gwh      = total_gwh_baseline,
  scenario_label  = "Without sediment",
  fit_ntd_per_kwh = FIT_NTD_PER_KWH,
  ltv             = LTV,
  r_loan          = R_LOAN
)



```

### Finance — baseline with sediment adjustment

```{r}
## Finance — with sediment
fin_with_sed <- build_finance_outputs(
  annual_gwh      = gwh_with_sed,
  scenario_label  = "With sediment",
  fit_ntd_per_kwh = FIT_NTD_PER_KWH,
  ltv             = LTV,
  r_loan          = R_LOAN
)

# Sensitivity across rate × LTV grid
sens_no_sed <- run_sensitivity(
  annual_revenue_ntd = fin_no_sed$annual_revenue_ntd,
  r_loan_seq         = R_LOAN_RANGE,
  ltv_seq            = LTV_RANGE,
  scenario_label     = "Without sediment"
)

sens_with_sed <- run_sensitivity(
  annual_revenue_ntd = fin_with_sed$annual_revenue_ntd,
  r_loan_seq         = R_LOAN_RANGE,
  ltv_seq            = LTV_RANGE,
  scenario_label     = "With sediment"
)

sens_combined <- bind_rows(sens_no_sed, sens_with_sed)

# Mark user's chosen parameters on the plot
user_point <- data.frame(
  scenario   = c("Without sediment", "With sediment"),
  r_loan_pct = R_LOAN * 100,
  ltv_pct    = LTV * 100
)

plot_npv_feasibility_frontier(sens_combined) +
  geom_point(
    data   = user_point,
    aes(x = r_loan_pct, y = ltv_pct),
    shape  = 23,
    size   = 4,
    fill   = "white",
    colour = "black",
    stroke = 1.2
  ) +
  labs(caption = paste0(
    "Diamond = current parameter setting",
    " (LTV = ", LTV * 100, "%, rate = ", R_LOAN * 100, "%)\n",
    "CAPEX = NTD ", CAPEX_NTD / 1e9, "B; ",
    "FIT = NTD ", FIT_NTD_PER_KWH, "/kWh; ",
    "Equity rate = ", R_EQUITY * 100, "%"
  ))
```

## Immediate-dispatch reference

```{r}
sim_forward24 <- run_reservoir_simulation_forward24(
  daily_flow = daily_flow,
  e_flow_mode = "recommended",
  scale_w2 = TRUE
)

annual_forward24 <- summarise_annual_generation(sim_forward24)

annual_forward24 |>
  summarise(
    mean_energy_gwh = mean(energy_gwh_total, na.rm = TRUE),
    p10_energy_gwh = quantile(energy_gwh_total, 0.10, na.rm = TRUE),
    p90_energy_gwh = quantile(energy_gwh_total, 0.90, na.rm = TRUE),
    mean_revenue_b_ntd = mean(revenue_ntd, na.rm = TRUE) / 1e9,
    mean_spill_m3 = mean(spill_m3_W1 + spill_m3_W2, na.rm = TRUE)
  )

```

## Forecast-Informed Operation Window

```{r}
plot_operation_window(
  sim_df = sim_forward24$W1,
  start_date = "2010-08-01",
  end_date = "2010-08-15",
  title_text = "W1 forecast-informed operation during high-flow period"
)

```

## Immediate Dispatch Simulation

```{r}
sim_instant <- run_reservoir_simulation(
  daily_flow  = daily_flow,
  e_flow_mode = "recommended"
)

plot_operation_window(
  sim_df     = sim_instant$W1,
  start_date = "2010-08-01",
  end_date   = "2010-08-15",
  title_text = "W1 immediate-dispatch operation during high-flow period"
)

plot_operation_window(
  sim_df     = sim_instant$W2,
  start_date = "2010-08-01",
  end_date   = "2010-08-15",
  title_text = "W2 immediate-dispatch operation during high-flow period"
)

```

## Annual generation summary — immediate dispatch

```{r}
annual_instant <- sim_instant$combined |>
  mutate(year = lubridate::year(date)) |>
  group_by(year) |>
  summarise(
    energy_gwh_W1 = sum(energy_kwh_W1, na.rm = TRUE) / 1e6,
    energy_gwh_W2 = sum(energy_kwh_W2, na.rm = TRUE) / 1e6,
    energy_gwh_total = sum(energy_kwh_total, na.rm = TRUE) / 1e6,
    revenue_ntd = sum(energy_kwh_total, na.rm = TRUE) * .FIN$fit_ntd_per_kwh,
    .groups = "drop"
  )

```

## 20-year NPV trajectories

```{r}
npv_forward24_20yr <- run_npv_20yr_from_generation(
  annual_generation_df = annual_forward24,
  ltv                  = LTV,
  r_loan               = R_LOAN,
  discount_rate        = R_EQUITY
)

npv_forward24_20yr <- run_npv_20yr_from_generation(
  annual_generation_df = annual_forward24,
  ltv = 0.75,
  r_loan = 0.03,
  discount_rate = 0.08
)
  
npv_instant_20yr <- run_npv_20yr_from_generation(
  annual_generation_df = annual_instant,
  ltv                  = LTV,
  r_loan               = R_LOAN,
  discount_rate        = R_EQUITY
)

plot_npv_financial_line_20yr(
  npv_forward24_20yr,
  title_text = "Forward-looking operation: 20-year NPV trajectory"
)

plot_npv_financial_line_20yr(
  npv_instant_20yr,
  title_text = "Immediate-dispatch operation: 20-year NPV trajectory"
)

plot_npv_comparison_20yr(
  npv_instant   = npv_instant_20yr,
  npv_forward24 = npv_forward24_20yr,
  title_text    = "Immediate vs forecast-informed operation: 20-year NPV"
)
```

## Feasibility frontier — operations

```{r}

frontier_instant <- run_sensitivity(
  annual_revenue_ntd = mean(annual_instant$revenue_ntd, na.rm = TRUE),
  r_loan_seq         = seq(0.01, 0.08, by = 0.005),
  ltv_seq            = seq(0.40, 0.90, by = 0.05),
  scenario_label     = "Immediate dispatch"
)

frontier_forward24 <- run_sensitivity(
  annual_revenue_ntd = mean(annual_forward24$revenue_ntd, na.rm = TRUE),
  r_loan_seq         = seq(0.01, 0.08, by = 0.005),
  ltv_seq            = seq(0.40, 0.90, by = 0.05),
  scenario_label     = "24-hour forecast-informed"
)

frontier_compare <- bind_rows(frontier_instant, frontier_forward24)

plot_sensitivity_frontier(
  frontier_compare,
  title_text = "Immediate vs forecast-informed operation: feasibility frontier"
)
```

## 20-year NPV trajectories

```{r}
plot_npv_comparison_20yr(
  npv_instant = npv_instant_20yr,
  npv_forward24 = npv_forward24_20yr,
  title_text = "Immediate vs forecast-informed operation: 20-year NPV"
)
npv_forward24_20yr <- run_npv_20yr_from_generation(
  annual_generation_df = annual_forward24,
  ltv = 0.75,
  r_loan = 0.03,
  discount_rate = 0.08
)

plot_npv_financial_line_20yr(
  npv_forward24_20yr,
  title_text = "Forward-looking operation: 20-year NPV trajectory"
)
npv_instant_20yr <- run_npv_20yr_from_generation(
  annual_generation_df = annual_instant,
  ltv = 0.75,
  r_loan = 0.03,
  discount_rate = 0.08
)

plot_npv_financial_line_20yr(
  npv_instant_20yr,
  title_text = "Immediate-dispatch operation: 20-year NPV trajectory"
)
```

### Operations Feasibility Frontier - Comparing 

```{r}
frontier_instant <- run_sensitivity(
  annual_revenue_ntd = mean(annual_instant$revenue_ntd, na.rm = TRUE),
  r_loan_seq = seq(0.01, 0.08, by = 0.005),
  ltv_seq = seq(0.40, 0.90, by = 0.05),
  scenario_label = "Immediate dispatch"
)

frontier_forward24 <- run_sensitivity(
  annual_revenue_ntd = mean(annual_forward24$revenue_ntd, na.rm = TRUE),
  r_loan_seq = seq(0.01, 0.08, by = 0.005),
  ltv_seq = seq(0.40, 0.90, by = 0.05),
  scenario_label = "24-hour forecast-informed"
)

frontier_compare <- bind_rows(
  frontier_instant,
  frontier_forward24
)

plot_sensitivity_frontier(
  frontier_compare,
  title_text = "Immediate vs forecast-informed operation: feasibility frontier"
)
```
