# Hydrological Variability Assessment of Lishan Streamflow

## Notation and Data Structure

This analysis evaluates long-term hydrological variability using ten-day average streamflow records from the Lishan Hydrological Station.

**Notation**

- `cms`: Streamflow in cubic meters per second.
- `tendays`: Ten-day period index within each year.
- `EIA`: Original Environmental Impact Assessment reference period.
- `EIA Difference`: Environmental Impact Assessment Difference / amendment reference period.
- `Current`: Most recent comparison period.
- Each year contains 36 ten-day observations.
- Year columns are already stored in Gregorian / CE format in the CSV file, such as `1959`, `1960`, ..., `2019`.

**Period Definitions**

| Period Label | Year Range | Description |
|---|---:|---|
| `EIA` | 1959–1995 | Original EIA reference period |
| `EIA Difference` | 1969–2005 | EIA Difference / amendment reference period |
| `Current` | 1983–2019 | Recent comparison period |

## Packages

```{r}
#| message: false
#| warning: false

packages <- c(
  "tidyverse",
  "here",
  "car"
)

installed_packages <- rownames(installed.packages())

for (pkg in packages) {
  if (!pkg %in% installed_packages) {
    install.packages(pkg)
  }
}

invisible(lapply(packages, library, character.only = TRUE))
```

## Read Data

```{r}
#| message: false
#| warning: false

flow_data <- read_csv(
  here("data", "lishan_flow.csv")
)

glimpse(flow_data)
```

## Check Data Structure

```{r}
# Check number of ten-day observations
nrow(flow_data)

# Check number of columns
# Expected: 1 tendays column + year columns
ncol(flow_data)

# Extract year columns
year_cols <- names(flow_data)[-1]

# Check first and last year
year_cols[1]
year_cols[length(year_cols)]
```

## Annual Ten-Day Streamflow Distribution

```{r}
#| fig.width: 11
#| fig.height: 6
#| fig.cap: "Annual distribution of ten-day average streamflow at Lishan Hydrological Station. Boxes represent the within-year distribution of ten-day streamflow values. The red annotation indicates the EIA reference period, while the blue annotation indicates the EIA Difference reference period."

# Remove tendays column for plotting
flow_matrix <- flow_data[, -1]

# Convert year column names to numeric years
years <- as.numeric(year_cols)

# Main boxplot
boxplot(
  flow_matrix,
  outline = FALSE,
  xaxt = "n",
  yaxt = "n",
  xlab = "Year",
  ylab = "Ten-day Average Streamflow (cms)",
  main = "Hydrological Variability of Lishan Ten-day Average Streamflow",
  xlim = c(-1, length(years) + 1),
  ylim = c(0, 70),
  col = "grey90",
  border = "grey40"
)

# X-axis minor tick marks
axis(
  side = 1,
  at = seq_along(years),
  labels = rep("", length(years)),
  col = "grey70"
)

# X-axis major year labels every five years
axis(
  side = 1,
  at = seq(1, length(years), 5),
  labels = years[seq(1, length(years), 5)]
)

# Y-axis minor tick marks
axis(
  side = 2,
  at = seq(0, 70, 5),
  labels = rep("", length(seq(0, 70, 5))),
  col = "grey70"
)

# Y-axis major labels
axis(
  side = 2,
  at = seq(0, 70, 10),
  labels = seq(0, 70, 10),
  las = 1
)

# Add horizontal grid lines
abline(
  h = seq(0, 70, 10),
  col = "grey85",
  lty = "dotted"
)

# Re-draw boxplot above grid lines
boxplot(
  flow_matrix,
  outline = FALSE,
  xaxt = "n",
  yaxt = "n",
  add = TRUE,
  col = "grey90",
  border = "grey40"
)

# Annual median trend line
lines(
  x = seq_along(years),
  y = apply(flow_matrix, 2, median, na.rm = TRUE),
  col = "grey20",
  lwd = 3
)

# Period index positions
eia_start <- which(years == 1959)
eia_end   <- which(years == 1995)

eia_diff_start <- which(years == 1969)
eia_diff_end   <- which(years == 2005)

current_start <- which(years == 1983)
current_end   <- which(years == 2019)

# EIA annotation
text(
  x = mean(c(eia_start, eia_end)),
  y = 4,
  labels = "EIA\n1959–1995",
  col = "lightcoral",
  cex = 0.9
)

lines(
  x = c(eia_start, eia_end),
  y = c(1.5, 1.5),
  col = "lightcoral",
  lwd = 3
)

lines(
  x = c(eia_start, eia_start),
  y = c(0.5, 2.5),
  col = "lightcoral",
  lwd = 3
)

lines(
  x = c(eia_end, eia_end),
  y = c(0.5, 2.5),
  col = "lightcoral",
  lwd = 3
)

# EIA Difference annotation
text(
  x = mean(c(eia_diff_start, eia_diff_end)),
  y = 8,
  labels = "EIA Difference\n1969–2005",
  col = "royalblue",
  cex = 0.9
)

lines(
  x = c(eia_diff_start, eia_diff_end),
  y = c(5.5, 5.5),
  col = "royalblue",
  lwd = 3
)

lines(
  x = c(eia_diff_start, eia_diff_start),
  y = c(4.5, 6.5),
  col = "royalblue",
  lwd = 3
)

lines(
  x = c(eia_diff_end, eia_diff_end),
  y = c(4.5, 6.5),
  col = "royalblue",
  lwd = 3
)

# Optional current period annotation
text(
  x = mean(c(current_start, current_end)),
  y = 12,
  labels = "Current\n1983–2019",
  col = "darkgreen",
  cex = 0.9
)

lines(
  x = c(current_start, current_end),
  y = c(9.5, 9.5),
  col = "darkgreen",
  lwd = 3
)

lines(
  x = c(current_start, current_start),
  y = c(8.5, 10.5),
  col = "darkgreen",
  lwd = 3
)

lines(
  x = c(current_end, current_end),
  y = c(8.5, 10.5),
  col = "darkgreen",
  lwd = 3
)

# Legend
legend(
  "topright",
  legend = c(
    "Annual median",
    "EIA: 1959–1995",
    "EIA Difference: 1969–2005",
    "Current: 1983–2019"
  ),
  col = c("grey20", "lightcoral", "royalblue", "darkgreen"),
  lwd = c(3, 3, 3, 3),
  bty = "n"
)
```

```{r}
#| fig.width: 11
#| fig.height: 6
#| fig.cap: "Annual distribution of ten-day average streamflow at Lishan Hydrological Station. Boxes represent the within-year distribution of ten-day streamflow values. Colored annotations at the bottom indicate the EIA, EIA Difference, and Current reference periods."

# Slightly enlarge left and bottom margins for annotations
par(mar = c(5, 6, 4, 2) + 0.1)

# Remove tendays column for plotting
flow_matrix <- flow_data[, -1]

# Convert year column names to numeric years
years <- as.numeric(year_cols)

# Main boxplot
boxplot(
  flow_matrix,
  outline = FALSE,
  xaxt = "n",
  yaxt = "n",
  xlab = "Year",
  ylab = "Ten-day Average Streamflow (cms)",
  main = "Hydrological Variability of Lishan Ten-day Average Streamflow",
  xlim = c(-1, length(years) + 1),
  ylim = c(0, 70),
  col = "grey90",
  border = "grey40"
)

# X-axis minor tick marks
axis(
  side = 1,
  at = seq_along(years),
  labels = rep("", length(years)),
  col = "grey70"
)

# X-axis major year labels every five years
axis(
  side = 1,
  at = seq(1, length(years), 5),
  labels = years[seq(1, length(years), 5)]
)

# Y-axis minor tick marks
axis(
  side = 2,
  at = seq(0, 70, 5),
  labels = rep("", length(seq(0, 70, 5))),
  col = "grey70"
)

# Y-axis major labels
axis(
  side = 2,
  at = seq(0, 70, 10),
  labels = seq(0, 70, 10),
  las = 1
)

# Add horizontal grid lines
abline(
  h = seq(0, 70, 10),
  col = "grey85",
  lty = "dotted"
)

# Re-draw boxplot above grid lines
boxplot(
  flow_matrix,
  outline = FALSE,
  xaxt = "n",
  yaxt = "n",
  add = TRUE,
  col = "grey90",
  border = "grey40"
)

# Annual median trend line
lines(
  x = seq_along(years),
  y = apply(flow_matrix, 2, median, na.rm = TRUE),
  col = "grey20",
  lwd = 3
)

# Period index positions
eia_start <- which(years == 1959)
eia_end   <- which(years == 1995)

eia_diff_start <- which(years == 1969)
eia_diff_end   <- which(years == 2005)

current_start <- which(years == 1983)
current_end   <- which(years == 2019)

# Bottom annotation y positions
y_eia      <- 2
y_eia_diff <- 5
y_current  <- 8

# Left-side labels
# text(x = 0.3, y = y_eia, labels = "EIA\n1959–1995", col = "lightcoral", cex = 0.9, adj = c(0, 0.5) )

# text( x = 0.3, y = y_eia_diff, labels = "EIA Difference\n1969–2005", col = "royalblue", cex = 0.9, adj = c(0, 0.5) )

# text( x = 0.3, y = y_current, labels = "Current\n1983–2019", col = "darkgreen", cex = 0.9, adj = c(0, 0.5) )

# EIA annotation line
lines(
  x = c(eia_start, eia_end),
  y = c(y_eia, y_eia),
  col = "lightcoral",
  lwd = 3
)

lines(
  x = c(eia_start, eia_start),
  y = c(y_eia - 0.8, y_eia + 0.8),
  col = "lightcoral",
  lwd = 3
)

lines(
  x = c(eia_end, eia_end),
  y = c(y_eia - 0.8, y_eia + 0.8),
  col = "lightcoral",
  lwd = 3
)

# EIA Difference annotation line
lines(
  x = c(eia_diff_start, eia_diff_end),
  y = c(y_eia_diff, y_eia_diff),
  col = "royalblue",
  lwd = 3
)

lines(
  x = c(eia_diff_start, eia_diff_start),
  y = c(y_eia_diff - 0.8, y_eia_diff + 0.8),
  col = "royalblue",
  lwd = 3
)

lines(
  x = c(eia_diff_end, eia_diff_end),
  y = c(y_eia_diff - 0.8, y_eia_diff + 0.8),
  col = "royalblue",
  lwd = 3
)

# Current annotation line
lines(
  x = c(current_start, current_end),
  y = c(y_current, y_current),
  col = "darkgreen",
  lwd = 3
)

lines(
  x = c(current_start, current_start),
  y = c(y_current - 0.8, y_current + 0.8),
  col = "darkgreen",
  lwd = 3
)

lines(
  x = c(current_end, current_end),
  y = c(y_current - 0.8, y_current + 0.8),
  col = "darkgreen",
  lwd = 3
)

# Legend
legend(
  "topright",
  legend = c(
    "Annual median",
    "EIA: 1959–1995",
    "EIA Difference: 1969–2005",
    "Current: 1983–2019"
  ),
  col = c("grey20", "lightcoral", "royalblue", "darkgreen"),
  lwd = c(3, 3, 3, 3),
  bty = "n"
)
```



## Define Comparison Periods

```{r}
# The CSV year columns are assumed to be already coded as Gregorian / CE years.

eia <- flow_data |>
  select(tendays, `1959`:`1995`) |>
  select(-tendays) |>
  unlist(use.names = FALSE)

eia_difference <- flow_data |>
  select(tendays, `1969`:`2005`) |>
  select(-tendays) |>
  unlist(use.names = FALSE)

current <- flow_data |>
  select(tendays, `1983`:`2019`) |>
  select(-tendays) |>
  unlist(use.names = FALSE)

length(eia)
length(eia_difference)
length(current)
```

## Build Period Comparison Dataset

```{r}
flow_period_df <- tibble(
  cms = c(eia, eia_difference, current),
  period = rep(
    c(
      "EIA: 1959–1995",
      "EIA Difference: 1969–2005",
      "Current: 1983–2019"
    ),
    each = length(eia)
  )
)

flow_period_df
```

## Summary Statistics by Period

```{r}
flow_period_summary <- flow_period_df |>
  group_by(period) |>
  summarise(
    n = n(),
    mean_cms = mean(cms, na.rm = TRUE),
    median_cms = median(cms, na.rm = TRUE),
    sd_cms = sd(cms, na.rm = TRUE),
    variance_cms = var(cms, na.rm = TRUE),
    q03_cms = quantile(cms, 0.03, na.rm = TRUE),
    q25_cms = quantile(cms, 0.25, na.rm = TRUE),
    q75_cms = quantile(cms, 0.75, na.rm = TRUE),
    max_cms = max(cms, na.rm = TRUE),
    .groups = "drop"
  )

flow_period_summary
```

## Mean Difference Tests

```{r}
# Mean comparison between EIA and EIA Difference
t.test(
  eia,
  eia_difference,
  var.equal = TRUE
)

# Mean comparison between EIA Difference and Current
t.test(
  eia_difference,
  current,
  var.equal = TRUE
)
```

## Variance Difference Tests

```{r}
# Variance comparison between EIA and EIA Difference
var.test(
  eia,
  eia_difference,
  alternative = "two.sided"
)

# Variance comparison between EIA Difference and Current
var.test(
  eia_difference,
  current,
  alternative = "two.sided"
)
```

## Non-parametric Distributional Comparison

```{r}
# Wilcoxon rank-sum test between EIA Difference and Current

wilcox.test(
  eia_difference,
  current,
  alternative = "two.sided",
  mu = 0,
  exact = FALSE
)
```

## One-way ANOVA

```{r}
aov_cms <- aov(
  cms ~ period,
  data = flow_period_df
)

summary(aov_cms)
```

## Period-level Boxplot

```{r}
#| fig.width: 8
#| fig.height: 5
#| fig.cap: "Comparison of ten-day average streamflow distributions across the EIA, EIA Difference, and Current periods."

boxplot(
  cms ~ period,
  data = flow_period_df,
  outline = FALSE,
  xlab = "Reference Period",
  ylab = "Ten-day Average Streamflow (cms)",
  main = "Comparison of Lishan Streamflow Across Reference Periods",
  col = "grey90",
  border = "grey40"
)
```

## Levene Test for Equality of Variance

```{r}
leveneTest(
  cms ~ period,
  data = flow_period_df,
  center = mean
)
```

## Low-flow Threshold

```{r}
# Estimate the 3rd percentile low-flow threshold across all year columns.
# The tendays column is excluded.

low_flow_q03 <- quantile(
  unlist(flow_data[, -1], use.names = FALSE),
  0.03,
  na.rm = TRUE
)

low_flow_q03
```

## Interpretation Notes

The analysis compares whether the streamflow distribution differs across three reference periods: the original EIA period, the EIA Difference period, and the Current period. The boxplot shows changes in annual ten-day streamflow distribution, while the statistical tests evaluate whether the mean, variance, and broader distributional characteristics differ across periods.

The t-tests evaluate mean differences. The F-tests and Levene test evaluate variance differences. The Wilcoxon test provides a non-parametric comparison that is less dependent on normality assumptions. The 3rd percentile threshold provides a low-flow reference value for assessing hydrological stress or drought-sensitive flow conditions.